---
title: "Hierarchical Agglomeration clustering.rmd"
author: "Ke Li"
date: "2018/11/27"
output: html_document
highlight: tango 
---

## 1. Data Summary 
\
Load all the libraries we need for our analysis:
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(cluster)
library(fpc)
library(ggplot2)
library(tidyr)
library(dplyr)
library(data.table)
library(psych)
library(cowplot)

```

We need to load data into R by using read.csv since it's in csv format.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# set working directory
# setwd("~/Desktop/group_proj")

# import data
data=read.csv("https://raw.githubusercontent.com/Anranmg/project506/master/BlackFriday.csv",header=TRUE, sep=',')
```


Product_category 2 and product_category 3 are products belong to other catagories, it recalculate same product, thus, we drop these two product_categories.\
By manipulating data, we get a new dataset with user_id, occupation, age, gender and ratio in 18 categories for each user.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# profiling attributes: Occupation,Age,Gender
# segementation attributes: ratio_in_cate1-ratio_in_cate18
# Since all segementation variables are on a similar scale, we skip the feature scaling step.
data = data%>%
  select(User_ID,Occupation,Age,Gender,Product_Category_1,Purchase)%>%
  group_by(User_ID)%>%
  mutate(total=sum(Purchase))%>%
  group_by(User_ID,Product_Category_1,Occupation,Age,Gender)%>%
  summarise(Purchase = sum(Purchase), total = mean(total))%>%
  mutate(Purchase=Purchase/total)%>%
  spread(Product_Category_1,Purchase)%>%
  replace(.,is.na(.),0)%>%
  select(-total)

names(data)[5:22]=paste0("ratio_in_cate",1:18,sep="")

```


Simple summary statistics of our dataset
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
psych::describe(data[,2:22])%>%
  select(vars, n, mean, sd, median, min, max, range)

```

check structure and list first 5 obs
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
head(data,5)
```

Choosing ratio in categories as segementaion variables, and using euclidean distance measure to calculate pairwise distance
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data1=data[5:22]
m=dist(data1)%>%
  as.matrix()
```

Show pairwise distance between first 5 consumers
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
m[1:5,1:5]
```


To test whether the method is fit for this data.\
Select histogram of ratio_in_cate1 and ratio_in_cate5 \

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot1=ggplot(data,aes(x=ratio_in_cate1))+
  geom_histogram(col="black",fill="white")+
  xlab("histogram of ratio_in_cate1")+
  ylab("frequency")
plot2=ggplot(data,aes(x=ratio_in_cate5))+
  geom_histogram(col="black",fill="white")+
  xlab("histogram of ratio_in_cate5")+
  ylab("frequency")
cowplot::plot_grid(plot1,plot2)
```
\
These two histograms show that customer have different preference on different category products.\
We could find 'mountains and valleys' patterns in these two graphs, indicating being potential segements for later analysis. \


## 2. Clustering analysis
\
Using hierarchical clustering method, and get the dendrogram.\
Method: Agglomeration method \
Linkage: ward \
Use euclidean method to calculate distance.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
sol = hclust(dist(data1), method = "ward.D")
plot(sol,cex = 0.3,hang = - 10, main = "Ward-linkage Cluster Dendrogram")
rect.hclust(sol, k=20, border="red")

```

*Select number of clusters* \
Based on ch index

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
stat=data.frame(no.=c(2:15))
fun=function(k){
  cluster_id = cutree(as.hclust(sol), k)
  return(calinhara(data1,cluster_id))
}
stat$calinhara=sapply(stat$no.,fun)
stat
```
When clusters number is 2, calinhara has max value.\
Thus, choose 2 be the number of clusters.\
\

*Clustering result*
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
cluster_id = cutree(as.hclust(sol), k = 2)
```

*Check the number of each cluster*
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
table(cluster_id)
data_cluster = data.frame(cluster_id)
tmp=cbind(cluster_id=data_cluster$cluster_id,data)
```

*clustering result in first 10 obs*
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
tmp[1:10,c(2,1)]
```

## 3. Profile and interpret the segments
#### Plot of Purchase Proportion
The plot shows average purchase proportion for each group and mean ourchase proportion for whole population.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
tb=as.data.table(tmp)
av=data.frame(cluster_id=0)%>%
  cbind(as.data.frame(tb[,lapply(.SD,mean),.SDcols=ratio_in_cate1:ratio_in_cate18]))%>%
  bind_rows(as.data.frame(tb[,lapply(.SD,mean),by=cluster_id,.SDcols=ratio_in_cate1:ratio_in_cate18]))%>%
  gather(key=type,value=measurement,-cluster_id)%>%
  mutate(cluster_id=factor(cluster_id),category=rep(1:18,each=3))

#table of average values within clusters
tb1=av%>%
  spread(key=cluster_id,val=measurement)
names(tb1)[3:5]=c("population","group1","group2")

#line plot of average values within clusters
av%>%
  ggplot(aes(x=category,y=measurement,group=cluster_id))+
  geom_line(aes(color=cluster_id))+
  geom_point(aes(color=cluster_id))+
  labs(title="Plot of Purchase Proportion",y="mean of cluster")+
  scale_color_manual(labels=c("Population","Group 1","Group 2"),
                     values=c("#999999", "#E69F00", "#56B4E9"))
```

#### Plot of Occupation Proportion
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# line plot of occupation
ocp=tmp%>%
  group_by(cluster_id)%>%
  mutate(n=1/n())%>%
  group_by(cluster_id,Occupation)%>%
  summarise(pct=sum(n))%>%
  bind_rows(tmp%>%
              mutate(n=1/n())%>%
              group_by(Occupation)%>%
              summarise(pct=sum(n))%>%
              mutate(cluster_id=0))
ocp$cluster_id=factor(ocp$cluster_id)

ocp%>%
  ggplot(aes(x=Occupation,pct,group=cluster_id))+
  geom_line(alpha=.5, aes(color=cluster_id))+
  geom_point(alpha=.5, aes(color=cluster_id))+
  labs(title="Plot of Occupation Proportion",y="mean of cluster")+
  scale_color_manual(labels=c("Population","Group 1","Group 2"),
                     values=c("#999999", "#E69F00", "#56B4E9"))
```

#### Distribution plot of age
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
tmp$cluster_id=factor(tmp$cluster_id)
ggplot(tmp,aes(x=Age,fill=cluster_id))+
  geom_histogram(alpha=.5, position="dodge",stat="count")+
  scale_color_discrete(labels=c("Group 1","Group 2"))+
  labs(title="Plot of Age Distribution")
```
\
According to the plot, we can see that the age is similar distributed in each group.\


*** From plot of Purchase Proportion and plot of Occupation Proportion, we can see group 2 have more people with occupation 2 and 20 than group 1, while group 2 has obvious purchase preference on product category 5 and 8. Thus, we can infer that people with occupation 2 and 20 may prefer buying product category 5 and 8.***


*** Group 1 has more people with occupation 13 and 17, while group 1 has large purchase ratio on product category 1, we can also speculate that people with occupation 13 and 17 have ralatively purchase preference on product category 1.***




